Capstone Movielens Report

### 2025-11-27 reports/capstone-movielens.site/02-3.OMR-model.Rmd

  - [naive_model_MSEs](#func.naive_model_MSEs) (described in section [OMR Model Helper Functions]) to compute _Mean Squared Errors_ using _5-Fold Cross-Validation_ method;

@@

::: {.noteblock data-latex=""}
In the code snippet above, we use the function [naive_model_RMSE](#func.naive_model_RMSE) described in the [OMR Model Helper Functions] section of [Appendix A: Support Functions] to calculate *RMSEs* for the sequence of the deviation values.
:::

>>>>

  - [naive_model_MSEs](#func.naive_model_MSEs) (described in section [OMR Model: Helper Functions]) to compute _Mean Squared Errors_ using _5-Fold Cross-Validation_ method;

@@

::: {.noteblock data-latex=""}
In the code snippet above, we use the function [naive_model_RMSE](#func.naive_model_RMSE) described in the [OMR Model: Helper Functions] section of [Appendix A: Support Functions] to calculate *RMSEs* for the sequence of the deviation values.
:::

### 2025-11-27 reports/capstone-movielens.site/02.5.UME-model.Rmd

To prove this fact, we can find out the movies with the highest number of ratings using the following code:
```{r, eval=FALSE}
edx.ordered_movie_ratings <- edx |> group_by(movieId, title) |>
  summarize(number_of_ratings = n()) |>
  arrange(desc(number_of_ratings))
```
```{r }
print(head(edx.ordered_movie_ratings))
```

Now, we can figure out the most given ratings in order from most to least:
```{r, eval=FALSE}
edx.rating_groups <- edx |>  group_by(rating) |>
     summarise(count = n()) |>
     arrange(desc(count))
```
```{r, include=TRUE}
print(edx.rating_groups)
```

@@

The following code allows us to summarize that in general, half-star ratings are less common than whole-star ratings (e.g., there are fewer ratings of 3.5 than there are ratings of 3 or 4, etc.):
```{r, include=TRUE}
print(edx.rating_groups |> arrange(rating))
```

>>>>

To prove this fact, we can find out the movies with the highest number of ratings using the following code:
```{r, eval=FALSE}
edx.ordered_movie_ratings <- edx |> group_by(movieId, title) |>
  summarize(number_of_ratings = n()) |>
  arrange(desc(number_of_ratings))
```
```{r }
head(edx.ordered_movie_ratings)
```

Now, we can figure out the most given ratings in order from most to least:
```{r, eval=FALSE}
edx.rating_groups <- edx |>  group_by(rating) |>
     summarise(count = n()) |>
     arrange(desc(count))
```
```{r, include=TRUE}
edx.rating_groups
```

@@

The following code allows us to summarize that in general, half-star ratings are less common than whole-star ratings (e.g., there are fewer ratings of 3.5 than there are ratings of 3 or 4, etc.):
```{r, include=TRUE}
edx.rating_groups |> arrange(rating)
```

### 2025-11-27 reports/capstone-movielens.site/02.6.UMGE-model.Rmd

As mentioned in [Chapter 24: *Regularization*, Section 24.4: *Exercises*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#exercises) (*Exercises 9-15*) of the *Course Textbook* the `Movielens` dataset also has a genres column. This column includes every genre that applies to the movie (some movies fall under several genres)[@IDS2_24-4].

@@

dim(genre_ratins_grp)
genre_ratins_grp_sorted <- genre_ratins_grp |> sort_by.data.frame(~ rating_avg)
print(genre_ratins_grp_sorted)

# Creating plot:
genre_ratins_grp |> 
  ggplot(aes(x = genres, y = rating_avg, ymin = rating_avg - 2*se, ymax = rating_avg + 2*se)) + 
  geom_point() +
  geom_errorbar() + 
  ggtitle("Average rating per Genre") +
  ylab("Average rating") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
Below are worst and best ratings categories:
```{r, echo=FALSE}
sprintf("The worst ratings are for the genre category: %s",
        genre_ratins_grp$genres[which.min(genre_ratins_grp$genres)])

@@

### UMGE Model: Support Functions
\

::: {.noteblock data-latex=""}
Some of the functions described in this section accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call as well, which we will also need for the *Regularization* techniques to apply to our models (for details, see the  [mean_reg](#func.mean_reg) function description in the Section [Common Helper Functions: Regularization](#appndx_a.CHF.regularization) section of [Appendix A: Support Functions]).
We have already explained that in the [UME Model Regularization] section. 
For now, we omit the `lambda` parameter, accepting its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

#### [train_user_movie_genre_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R#L40) Function {#func.train_user_movie_genre_effect}
\

We use this function to build and train our model using a _Train Set_ (passing it to the `train.sgr` parameter), which can be any sample of the `edx.sgr` dataset (or the entire dataset itself) described above in the [`edx.sgr` Object] section. Here, I only remind that it is the dataset created from the `edx` one by splitting its rows to ensure each belongs to a single _genre_. Here is the source code of the function:
```{r eval=FALSE}
train_user_movie_genre_effect <- function(train.sgr, lambda = 0){
  if (is.na(lambda)) {
    stop("Function: train_user_movie_genre_effect
`lambda` is `NA`")
  }

  genre_bias <- train.sgr |>
      left_join(edx.user_effect, by = "userId") |>
      left_join(rglr.UM_effect, by = "movieId") |>
      mutate(resid = rating - (mu + a + b)) |>
      group_by(genres) |>
      summarise(g = mean_reg(resid, lambda), n = n())

    train.sgr |>
      left_join(genre_bias, by = "genres") |>
      left_join(rglr.UM_effect, by = "movieId") |>
      group_by(movieId) |>
      summarise(g = mean(g))
}
```

#### [train_user_movie_genre_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R#L59) Function {#func.train_user_movie_genre_effect.cv}
\

We use this function, defined in the [UMG-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R#L59) script, to train the current model using the _5-Fold Cross Validation_ method. Below, we provide a slightly simplified version of the source code of that function:
```{r eval=FALSE}
train_user_movie_genre_effect.cv <- function(lambda = 0){
  # ...

put_log1("Function `train_user_movie_genre_effect.cv`:
Computing User+Movie+Genre Effects list for %1-Fold Cross Validation samples...", 
           CVFolds_N)

  start <- put_start_date()
  user_movie_genre_effects_ls <- lapply(kfold_index, function(fold_i){
    cv_fold_dat <- edx_CV[[fold_i]]

    put_log2("Processing User+Movie+Genre Effects for %1-Fold Cross Validation samples (Fold %2)...",
             CVFolds_N,
             fold_i)
    umg_effect <- cv_fold_dat$train.sgr |> train_user_movie_genre_effect(lambda)

    put_log2("User+Movie+Genre Effects have been computed for the Fold %1 
of the %2-Fold Cross Validation samples.",
             fold_i,
             CVFolds_N)
    umg_effect
  })
  put_end_date(start)
  put_log1("Function `train_user_movie_genre_effect.cv`:
User+Movie+Genre Effects list has been computed for %1-Fold Cross Validation samples.", 
           CVFolds_N)

  user_movie_genre_effects_united <- union_cv_results(user_movie_genre_effects_ls)

  user_movie_genre_effect <- user_movie_genre_effects_united |>
    group_by(movieId) |>
    summarise(g = mean(g))

  if(lambda == 0) put_log("Function `train_user_movie_genre_effect.cv`:
Training completed: User+Movie+Genre Effects model.")
  else put_log1("Function `train_user_movie_genre_effect.cv`:
Training completed: User+Movie+Genre Effects model for lambda: %1...",
                lambda)

  user_movie_genre_effect
}
```

::: {.noteblock data-latex=""}
Here we use the function call [union_cv_results](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L432) to aggregate the _5-Fold Cross Validation_ method results (for details, see the [union_cv_results](#func.union_cv_results) function description in the [Data Helper Functions](#appndx_a.data_helper.functions) section of [Appendix A: Support Functions]).
:::

#### `calc_user_movie_genre_effect_MSE` Function
\

The source code of the function [calc_user_movie_genre_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R#L110) defined in the [UMG-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R) script to calculate the _Mean Squared Error (MSE)_ of the _UMGE Model_ for the given _Test Set_ is provided below:
```{r eval=FALSE}
calc_user_movie_genre_effect_MSE <- function(test_set, umg_effect){
  test_set |>
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(umg_effect, by = "movieId") |>
    mutate(resid = rating - clamp(mu + a + b + g)) |> 
    pull(resid) |> mse()
}
```

#### `calc_user_movie_genre_effect_MSE.cv` Function
\

The source code of the function [calc_user_movie_genre_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R#L122) defined in the [UMG-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R) script to calculate the _5-Fold Cross Validation MSE_ result of the _UMGE Model_ is provided below:
```{r eval=FALSE}
calc_user_movie_genre_effect_MSE.cv <- function(umg_effect){
  put_log("Computing RMSEs.ResultTibble on Validation Sets...")
  start <- put_start_date()
  user_movie_genre_effects_MSEs <- sapply(edx_CV, function(cv_dat){
    cv_dat$validation_set |> calc_user_movie_genre_effect_MSE(umg_effect)
  })
  put_end_date(start)

  plot(user_movie_genre_effects_MSEs)
  put_log1("MSE values have been plotted for the %1-Fold Cross Validation samples.", 
           CVFolds_N)

  mean(user_movie_genre_effects_MSEs)
}
```

#### `calc_user_movie_genre_effect_RMSE` Function
\

The source code of the function [calc_user_movie_genre_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R#L118) defined in the [UMG-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R) script to calculate the _Root Mean Squared Error (RMSE)_ of the _UMGE Model_ for the given _Test Set_ is provided below:

```{r eval=FALSE}
calc_user_movie_genre_effect_RMSE <- function(test_set, umg_effect){
  umg_mse <- test_set |> calc_user_movie_genre_effect_MSE(umg_effect)
  sqrt(umg_mse)
}
```

#### `calc_user_movie_genre_effect_RMSE.cv` Function { #func.calc_user_movie_genre_effect_RMSE.cv }
\

The source code of the function [calc_user_movie_genre_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R#L136) defined in the [UMG-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMG-effect.functions.R) script to calculate the _5-Fold Cross Validation **RMSE**_ result of the _UMGE Model_ is provided below:

```{r eval=FALSE}
calc_user_movie_genre_effect_RMSE.cv <- function(umg_effect){
  umg_effect_RMSE <- sqrt(calc_user_movie_genre_effect_MSE.cv(umg_effect))
  put_log2("%1-Fold Cross Validation ultimate RMSE: %2", 
           CVFolds_N, 
           umg_effect_RMSE)

  umg_effect_RMSE
}
```

>>>>

As we can see from the `edx` dataset structure, the `Movielens` dataset also has a genres column. This column includes every genre that applies to the movie (some movies fall under several genres):
```{r}
str(edx)
```

@@

# Creating plot:
genre_ratins_grp |> 
  ggplot(aes(x = genres, y = rating_avg, ymin = rating_avg - 2*se, ymax = rating_avg + 2*se)) + 
  geom_point() +
  geom_errorbar() + 
  ggtitle("Average rating per Genre") +
  ylab("Average rating") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Below are worst and best ratings categories:
```{r, echo=FALSE}
sprintf("The worst ratings are for the genre category: %s",
        genre_ratins_grp$genres[which.min(genre_ratins_grp$genres)])

@@

### 2025-11-27 reports/capstone-movielens.site/05.appendix.Rmd

#### [splitGenreRows](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L73) Function {#func.splitGenreRows}
\
Splits rows of the original dataset passed in the [data](#) argument for movies that belong to multiple genres.

@@

```{r eval=FALSE}
dat.sgr <- splitGenreRows(data)
```
:::

@@

::: { #func.splitGenreRows.details .sidebar }
##### Details
\

This function is used internally in the [make_source_datasets](#func.make_source_datasets) function described above in the [Initializing Input Datasets] section to create the `edx.sgr` dataset described below in the [`edx.sgr` Object] section of the [Appendix B: Models Training Datasets].

@@

@@

The source code of the [splitGenreRows](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L73) function is shown below:
```{r eval=FALSE}
```

@@

Trains (and optionally *regularizes* if the [lambda](#) argument is non-default) the *User+Movie Effect (UME) Model* using the training dataset passed in the [train_set](#func.train_user_movie_effect.args.train_set) argument.

@@

Trains (and optionally *regularizes* if the [lambda](#) argument is non-default) the *User+Movie Effect (UME) Model* using the training dataset passed in the [train_set](#func.train_user_movie_genre_effect.args.train_set) argument.

@@

@@

::: { #func.train_user_movie_genre_effect.args.train_set .sidebar }
  - **train_set:** Training dataset;

@@

::: { #func.train_user_movie_genre_effect.details .sidebar }
##### Details
\

The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_genre_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::

@@

Data frame representing the trained (and also *regularized* if [lambda](#func.train_user_movie_genre_effect.args.lambda) arg value other than default) *UM Effect Model*.
:::

@@

@@

Data frame representing the *UM Effect Model* trained (and also *regularized* if the [lambda](#func.train_user_movie_genre_effect.cv.args.lambda) argument value is other than the default) with use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object]..
:::

@@

@@

:::
::: { #func.calc_user_movie_genre_effect_MSE.args.um_effect .sidebar }
  - **um_effect:** Data frame representing the *UM Effect Model*.

@@

The function calculates *Mean Squared Error (MSE)* for the *UME Model* passed in the [um_effect](#func.calc_user_movie_genre_effect_MSE.args.um_effect) argument using the validation dataset passed in the [test_set](#func.calc_user_movie_genre_effect_MSE.args.test_set) arg.

@@

@@

@@

::: { #func.calc_user_movie_genre_effect_MSE.cv.args.um_effect .sidebar }
  - **um_effect:** Data frame representing the *UM Effect Model*.

@@

The function calculates *MSE* for the *UME Model* passed in the [um_effect](#func.calc_user_movie_genre_effect_MSE.cv.args.um_effect) argument using the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

@@

@@

::: { #func.calc_user_movie_genre_effect_RMSE.cv.args.um_effect .sidebar }
  - **um_effect:** Data frame representing the *UM Effect Model*.

@@

The function calculates *Root Squared Error (RMSE)* for the *UME Model* passed in the [um_effect](#func.calc_user_movie_genre_effect_RMSE.cv.args.um_effect) argument using the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

@@

@@

Validates the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMG_effect.cv.args.lambda) argument for the *User+Movie Effect (UME) Model* in the *regularization process* using *penalized approach* with use of *K-Fold Cross-Validation*.

@@

The function is used in the *regularization process* of the *UME Model* using the *penalized approach* to regularize predictions for the given *regularization parameter* $\lambda$ (passed in the [lambda](#func.regularize.test_lambda.UMG_effect.cv.args.lambda) argument) as explained in [Section 24.1 *Case study: recommendation systems / A penalized approach*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#a-penalized-approach) of the *Course Textbook (New Edition)*[@IDS2_24-1].

@@

The *Root Mean Squared Errors (RMSE)* value computed for the UME Model applying the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMG_effect.cv.args.lambda) argument.

>>>>

#### [splitGenreRows](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L73) Function {#func.splitGenreRows}
\
Splits rows of the original dataset passed in the [data](#func.splitGenreRows.args.data) argument for movies that belong to multiple genres.

@@

```{r eval=FALSE}
edx.sgr <- splitGenreRows(edx)
```
:::

@@

::: { #func.splitGenreRows.details .sidebar }
##### Details
\
This function is a wrapper for the [tidyr::separate_rows](https://www.rdocumentation.org/packages/tidyr/versions/1.3.0/topics/separate_rows) `R` function used in *This Project* to split rows belonging to multiple genres of a dataset containing a `genres` column, which is passed in the [data](#func.splitGenreRows.args.data) argument.

::: {.noteblock data-latex=""}
The `splitGenreRows` function is used internally in the [make_source_datasets](#func.make_source_datasets) function described above in the [Initializing Input Datasets] section to create the `edx.sgr` dataset described below in the [`edx.sgr` Object] section of the [Appendix B: Models Training Datasets].
:::
:::

@@

Dataset with the same structure as the one passed in the [data](#func.splitGenreRows.args.data) argument, but with rows that belong to multiple genres split.
:::

@@

The source code of the [splitGenreRows](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L73) function is shown below:
```{r eval=FALSE}
splitGenreRows <- function(data){
  put("Splitting dataset rows related to multiple genres...")
  start <- put_start_date()
  gs_splitted <- data |>
    separate_rows(genres, sep = "\\|")
  put("Dataset rows related to multiple genres have been splitted to have single genre per row.")
  put_end_date(start)
  gs_splitted
}
```

@@

Trains (and optionally *regularizes* if the [lambda](#func.train_user_movie_effect.args.lambda) argument value is non-default) the *User+Movie Effect (UME) Model* using the training dataset passed in the [train_set](#func.train_user_movie_effect.args.train_set) argument.

@@

Trains (and optionally *regularizes* if the [lambda](#func.train_user_movie_genre_effect.args.lambda) argument value is non-default) the *User+Movie+Genre Effect (UMGE) Model* using the training sample of the `Movielens` dataset (passed in the [train.sgr](#func.train_user_movie_genre_effect.args.train.sgr) argument), specially prepared to train the *UMGE Model*.

@@

```{r eval=FALSE}
  rglr.UMG_effect <- edx.sgr |> train_user_movie_genre_effect(UMGE.rglr.best_lambda)
```
:::

@@

::: { #func.train_user_movie_genre_effect.args.train.sgr .sidebar }
  - **train.sgr:** Training sample of the `Movielens` dataset with split rows belonging to multiple genres;

@@

::: { #func.train_user_movie_genre_effect.details .sidebar }
##### Details
\
The dataset passed in the [train.sgr](#func.train_user_movie_genre_effect.args.train.sgr) argument is assumed to be the result returned by the [splitGenreRows](#func.splitGenreRows) helper function described above in the *Data Helper Functions /* [Data Processing Functions] section of *This Appendix*.

For example, the [`edx.sgr` Object] is an appropriate dataset as the value of [train.sgr](#func.train_user_movie_genre_effect.args.train.sgr).

::: {.noteblock data-latex=""}
The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_genre_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

Data frame representing the trained (and also *regularized* if [lambda](#func.train_user_movie_genre_effect.args.lambda) arg value other than default) *UMG Effect Model*.
:::

@@

Trains the *User+Movie+Genre Effect (UMGE) Model* with use of *K-Fold Cross-Validation* as described in [Section 30.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_30-6].

@@

Data frame representing the *UMG Effect Model* trained (and also *regularized* if the [lambda](#func.train_user_movie_genre_effect.cv.args.lambda) argument value is other than the default) with use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object]..
:::

@@

Validates the *UMG Effect (UMGE) Model* produced by the  [train_user_movie_effect](#func.train_user_movie_effect) or [train_user_movie_effect.cv](#func.train_user_movie_effect.cv) function call.

@@

:::
::: { #func.calc_user_movie_genre_effect_MSE.args.um_effect .sidebar }
  - **um_effect:** Data frame representing the *UMG Effect Model*.

@@

The function calculates *Mean Squared Error (MSE)* for the *UMGE Model* passed in the [um_effect](#func.calc_user_movie_genre_effect_MSE.args.um_effect) argument using the validation dataset passed in the [test_set](#func.calc_user_movie_genre_effect_MSE.args.test_set) arg.

@@

*MSE* value calculated for the *UMGE Model*.
:::

@@

Validates the *UMG Effect (UMGE) Model* (produced by the [train_user_movie_effect](#func.train_user_movie_effect) or [train_user_movie_effect.cv](#func.train_user_movie_effect.cv) function call) with use of *K-Fold Cross-Validation* as described in [Section 30.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_30-6].

@@

::: { #func.calc_user_movie_genre_effect_MSE.cv.args.um_effect .sidebar }
  - **um_effect:** Data frame representing the *UMG Effect Model*.

@@

The function calculates *MSE* for the *UMGE Model* passed in the [um_effect](#func.calc_user_movie_genre_effect_MSE.cv.args.um_effect) argument using the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

@@

The *MSE* value calculated for the *UMG Effect Model* with use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object].
:::

@@

::: { #func.calc_user_movie_genre_effect_RMSE.cv.args.um_effect .sidebar }
  - **um_effect:** Data frame representing the *UMG Effect Model*.

@@

The function calculates *Root Squared Error (RMSE)* for the *UMGE Model* passed in the [um_effect](#func.calc_user_movie_genre_effect_RMSE.cv.args.um_effect) argument using the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

::: {.noteblock data-latex=""}
This function internally calls the [calc_user_movie_effect_MSE.cv](#func.calc_user_movie_effect_MSE.cv) function described above, to calculate the *MSE* for the *UMGE Model* with use of *K-Fold Cross-Validation*, and then calculates the square root of the result.

@@

The *RMSE* value calculated for the *UMG Effect Model* with use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object].
:::

@@

Validates the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMG_effect.cv.args.lambda) argument for the *User+Movie+Genre Effect (UMGE) Model* in the *regularization process* using *penalized approach* with use of *K-Fold Cross-Validation*.

@@

The function is used in the *regularization process* of the *UMGE Model* using the *penalized approach* to regularize predictions for the given *regularization parameter* $\lambda$ (passed in the [lambda](#func.regularize.test_lambda.UMG_effect.cv.args.lambda) argument) as explained in [Section 24.1 *Case study: recommendation systems / A penalized approach*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#a-penalized-approach) of the *Course Textbook (New Edition)*[@IDS2_24-1].

Internally it uses the [train_user_movie_effect.cv](#func.train_user_movie_effect.cv) and [calc_user_movie_effect_RMSE.cv](#func.calc_user_movie_effect_RMSE.cv) functions described above to validate the *UMGE Model* for the given $\lambda$ with use of the *K-Fold Cross-Validation* method, as described in [Section 30.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_30-6], where the value of `K` is determined by the length of the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets] (in our project we use `K = 5`).

@@

The *Root Mean Squared Errors (RMSE)* value computed for the UMGE Model applying the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMG_effect.cv.args.lambda) argument.
